A Unification of Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-14 colt-21/cvx-flats.tex
sec 6

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\relint}[1]{\mathrm{relint}(#1)}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\elic}{\mathrm{elic}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{cc\,dim}}
\newcommand{\codim}{\mathrm{codim}}
\newcommand{\Pcodim}{\mathcal{P}-\mathrm{codim}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\spn}{\mathrm{span}}
\newcommand{\range}{\mathrm{range}\,}
\newcommand{\zeros}[1]{\mathrm{ker}_\P\,#1}
\newcommand{\propdis}{\mu}
\newcommand{\affhull}{\mathrm{affhull}}
\newcommand{\epi}{\mathrm{epi}}
\newcommand{\cl}{\mathbf{cl}}
%\newcommand{\span}{\mathrm{span}}

@@

\begin{definition}[$\P$-codimension]
	The \emph{$\P$-codimension} of a convex set $F \subseteq \P$ is the minimum $d$ such that there exists a measurable function $V : \Y \to \reals^d$ such that $F = \zeros{V}$, and is denoted $\Pcodim(F)$.
\end{definition}

@@

\begin{corollary} \jessie{Indirect}
	Given target property $\gamma:\P \to\to \reals^d$, if there is some $p \in \P$ such that $|\gamma(p)| = 1$ with $\gamma(p) = \{r\}$, and there is not a flat $\zeros{\hat V_{u,p}} \subseteq \gamma_r$ of $\P$-codimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\hat \Gamma$.
\end{corollary}

\begin{corollary} \jessie{Indirect}
	Given target property $\gamma:\P \to\to \reals^d$, where $\P$ is defined over a finite set of outcomes $\Y$, if there is some $p \in \relint{\P}$ and there is not a flat $\zeros{\hat V_{u,p}} \subseteq \gamma_r$ of $\P$-codimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\gamma$.
\end{corollary}

@@

%commented out Jun 1 2020
%Theorem~\ref{thm:consistent-implies-indir-elic} allows us to use convex elicitation complexity as a tool to understand efficiency of consistent convex surrogates for a given property, which is more often what is given in a continuous estimation setting.
%For example, when one wants to learn an $\alpha$-quantile, we start with the property rather than a loss.
%In the literature, pinball loss $L(r,y) = (r-y)(\ones_{r \geq y} - \alpha)$ typically appears without explanation or justification as to why it is relevant in relation to the $\alpha$-quantile.
%Elicitation teaches us why pinball loss is consistent for learning a quantile as the pinball loss elicits the $\alpha$-quantile.
In continuous estimation problems, often one does not start with a target loss per se, but instead with a conditional statistic of the data one wishes to estimate, such as the conditional mean, variance, or entropy.
In this setting, Theorem~\ref{thm:cvx-flats} gives lower bounds on the prediction dimension of convex losses with a link to the desired conditional statistic, i.e., the convex elicitation complexity.
In particular,
% following an argument from~\citet{frongillo2018elicitation},
Theorem~\ref{thm:bayes-risk-lower-bound} below yields new bounds on the convex elicitation complexity of properties quantifying risk or uncertainty such as variance and entropy.
These bounds address an open question of~\citet{frongillo2018elicitation}, that of developing a theory of elicitation complexity for with respect to convex-elicitable properties.
Extending to infinite $\Y$, such as $\Y=\reals$, remains open.

>>>>

\newcommand{\reals}{\mathbb{R}}
\newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
\newcommand{\simplex}{\Delta_\Y}
\newcommand{\relint}[1]{\mathrm{relint}(#1)}
\newcommand{\prop}[1]{\mathrm{prop}[#1]}
\newcommand{\elic}{\mathrm{elic}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
\newcommand{\ccdim}{\mathrm{cc\,dim}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\spn}{\mathrm{span}}

\newcommand{\range}{\mathrm{range}\,}
\newcommand{\zeros}[1]{\mathrm{ker}_\P\,#1}
\newcommand{\codim}{\mathrm{codim}}
\newcommand{\Pcodim}{\mathcal{P}\!\text{-}\mathrm{codim}}
\newcommand{\Pcodimension}{$\mathcal{P}$-codimension}

\newcommand{\propdis}{\mu}
\newcommand{\affhull}{\mathrm{affhull}}
\newcommand{\epi}{\mathrm{epi}}
\newcommand{\cl}{\mathbf{cl}}
%\newcommand{\span}{\mathrm{span}}

@@

\begin{definition}[\Pcodimension]
	The \emph{\Pcodimension} of a convex set $F \subseteq \P$ is the minimum $d$ such that there exists a measurable function $V : \Y \to \reals^d$ such that $F = \zeros{V}$, and is denoted $\Pcodim(F)$.
\end{definition}

@@

\begin{corollary} \jessie{Indirect}
	Given target property $\gamma:\P \to\to \reals^d$, if there is some $p \in \P$ such that $|\gamma(p)| = 1$ with $\gamma(p) = \{r\}$, and there is not a flat $\zeros{\hat V_{u,p}} \subseteq \gamma_r$ of \Pcodimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\hat \Gamma$.
\end{corollary}

\begin{corollary} \jessie{Indirect}
	Given target property $\gamma:\P \to\to \reals^d$, where $\P$ is defined over a finite set of outcomes $\Y$, if there is some $p \in \relint{\P}$ and there is not a flat $\zeros{\hat V_{u,p}} \subseteq \gamma_r$ of \Pcodimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\gamma$.
\end{corollary}

@@

%commented out Jun 1 2020
%Theorem~\ref{thm:consistent-implies-indir-elic} allows us to use convex elicitation complexity as a tool to understand efficiency of consistent convex surrogates for a given property, which is more often what is given in a continuous estimation setting.
%For example, when one wants to learn an $\alpha$-quantile, we start with the property rather than a loss.
%In the literature, pinball loss $L(r,y) = (r-y)(\ones_{r \geq y} - \alpha)$ typically appears without explanation or justification as to why it is relevant in relation to the $\alpha$-quantile.
%Elicitation teaches us why pinball loss is consistent for learning a quantile as the pinball loss elicits the $\alpha$-quantile.
In continuous estimation problems, often one is not given a target loss, but instead a target (conditional) statistic of the data one wishes to estimate, such as the mean or variance.
In this setting, Theorem~\ref{thm:cvx-flats} gives lower bounds on the prediction dimension of convex losses with a link to the desired conditional statistic, i.e., the convex elicitation complexity.
In particular,
% following an argument from~\citet{frongillo2018elicitation},
Theorem~\ref{thm:bayes-risk-lower-bound} below yields new bounds on the convex elicitation complexity of statistics which quantify risk or uncertainty such as variance, entropy, or financial risk measures.

These bounds address an open question of~\citet{frongillo2020elicitation}, that of developing a theory of elicitation complexity for with respect to convex-elicitable properties.
The lower bounds of that work are essentially all with respect to identifiable properties, that is, properties $\gamma$ for which there is some $d$ where $\gamma_r$ is a flat of \Pcodimension $d$ for all $r\in\R$.
In contrast, properties elicited by non-smooth convex losses are generally not identifiable.
For example, the properties elicited by hinge loss and the abstain surrogate are not identifiable, as their level sets are not flats (see Figure~\ref{fig:fsd-bound}).
Establishing a theory of elicitation complexity for convex-elicitable properties therefore seemed to require entirely new ideas.
Surprisingly, we show that the central techniques of~\citet{frongillo2020elicitation} also apply to convex-elicitable properties, via \raf{MAIN THM}.
In particular, we can recover their main lower bound for the large class of Bayes risks.
